---
title: 'Lab: Class One Survey Analysis'
output:
  html_document:
    
---


### **INSTRUCTIONS**
####  **Complete all tasks 1-7 using reproducible coding practices (annotation of all code). Use R syntax to solve all problems (i.e. do not manually change values/variables in datasets or check them manually to make sure things are correct except as a double or triple check). Please cut and paste each problem before each R markdown chunk and annotate your code within the chunks.**


##### 1. Import class one survey data from our Github site (https://github.com/kijohnson/Advanced-Data-Analysis) (The dataset is called 'Class 1 Survey Fall 2026.csv'), calling the R dataframe that you create *C1survey*.
```{r}
C1survey <- read.csv("https://raw.githubusercontent.com/kijohnson/Advanced-Data-Analysis/refs/heads/main/Class%201%20Survey%20Fall%202026.csv")
```

##### 2. Determine: **a.** the number of observations (i.e. the number of people who filled out the survey) and **b.** the number of variables in the dataframe.    
```{r}
dim(C1survey)

nrow(C1survey)
ncol(C1survey)
```


##### 3. **a.** Rename the column variables to something shorter and that is descriptive of what the variable is about (for example *like_dogs* for the 'Do you like dogs?' question variable) and **b.** Display that they are renamed. **Hints**: You can use either the rename_with function or the names function. The list of names for time is:

```{r, echo=TRUE}
names(C1survey) <- c("id", "like_cats", "like_dogs", "have_desert", "slogan", "fav_day", "larkORowl", "fav_food", "fav_drink", "fav_season", "fav_month", "hobby", "program", "specialization", "stat_software", "R_exp", "coding_comfort", "coding_length", "top_three", "public_health_interest", "fav_num", "bday", "bmonth", "country", "state", "city",  "highest_educ_level")
```


```{r}
names(C1survey)
```

##### 4. Write code to determine and display the number of factor, integer, numerical, and character variables there are in the *C1survey* dataset. **Suggestion**: Try to ask ChatGPT for a solution if you are unsure. There is a really short answer to this but we want to see what students come up with!
```{r}
table(sapply(C1survey,class))
```

##### 5. **a.** Using code, check the *bday* and *bmonth* variables for any unusual or missing values (**hint**: use table function). If you find any, **b.** describe what you will do with the unusual values in a comment before or immediately following the code that does it, and **c.** after you have cleaned up any unusual values, find the median bday and month. *Note: please make sure that your bday and month variables are numeric before doing statistics
```{r}
names(C1survey)
class(C1survey$bday)
class(C1survey$bmonth)

table(C1survey$bday, useNA = "ifany")
table(C1survey$bmonth, useNA = "ifany")
```
```{r}
#The unusual values are written in words (not numbers). So I will recode them to a numeric value
C1survey$bday[C1survey$bday == "29-Nov"] <- "29"
C1survey$bday[C1survey$bday == "30-NOV"] <- "30"

C1survey$bmonth[C1survey$bmonth == "February"] <- "2"
C1survey$bmonth[C1survey$bmonth == "March"] <- "3"
C1survey$bmonth[C1survey$bmonth == "May"] <- "5"
C1survey$bmonth[C1survey$bmonth == "November"] <- "11"

C1survey$bday <- as.numeric(C1survey$bday)
C1survey$bmonth <- as.numeric(C1survey$bmonth)
```

```{r}
table(C1survey$bday, useNA = "ifany")
table(C1survey$bmonth, useNA = "ifany")

C1survey$bday[10] <- 30
C1survey$bmonth[22] <- 2

class(C1survey$bday)
class(C1survey$bmonth)
C1survey[is.na(C1survey$bday) | is.na(C1survey$bmonth),
         c("bday", "bmonth")]
median(C1survey$bday, na.rm = TRUE)
median(C1survey$bmonth, na.rm = TRUE)

```

##### 6. **a.** Create a new variable called *bseason* that gives the season according to Northern Meteorological season in which respondents were born (winter = 12/1 to 2/29, spring = 3/1 to 5/31, summer = 6/1 to 8/31, fall = 9/1 to 11/30). **b**. Using the table function, print a table of Seasons in the columns and bmonths in the rows that allows you to check that the coding is correct **c**. Sum the columns using the addmargins function to answer the question of how many classmates were born in each season? Include your answer in the code. 
```{r}
C1survey$bseason <- ifelse(C1survey$bmonth %in% c(12, 1, 2), "Winter",
                    ifelse(C1survey$bmonth %in% c(3, 4, 5), "Spring",
                    ifelse(C1survey$bmonth %in% c(6, 7, 8), "Summer",
                    ifelse(C1survey$bmonth %in% c(9, 10, 11), "Fall", NA))))

table(C1survey$bseason, useNA = "ifany")

table(C1survey$bmonth, C1survey$bseason)


addmargins(table(C1survey$bmonth, C1survey$bseason))

#Total number of participants born in the seasons are
 #Fall 15
 #Spring 5
 #Summer 5
 #Winter 5
```

##### 7. Pick your favorite variable or variables to analyze, come up with a question you want to answer with that variable or variables, generate the code, and provide an answer to your question. Describe what you did using comments (i.e. #'s).

```{r}
#I would like to know the most common hobby.

table(C1survey$hobby)

# I sorted the hobby counts in descending order.
sort(table(C1survey$hobby), decreasing = TRUE)

#The most common hobby reported was Reading (3 participants)

```






